\chapter{Conclusion}

Although the application developed during the course of this internship did not ultimately prove to be a viable solution to the problem at hand, the experience and knowledge aquired throughout this project have been invaluable. The study and research conducted on eBPF and kernel-based security provided a deeper understanding of recent technologies and their potential applications on the fields of monitoring and security.

The internship allowed for the exploration of cutting-edge technologies such as eBPF and its potential in enhancing security measures, thus expanding the understanding of these technologies as well as highlight the challenges and limitations of said technologies.

The development of this application in a company setting, proved itself to be a learning experience in of itself, providing an unique opportunity to interact with technologies that were not yet accustomed to. The cycle of development, from development, versioning and testing will surely prove to be an asset in future projects.

While the application developed may not be a bullet-proof solution, it demonstrates the potential of eBPF in enhancing security measures. 
The application has proven to be a functional tool that can contribute to the overall security of a system, even if in a limited way.
While the end result may not serve for this particular use case the study and research will surely prove quite useful when trying to leverage eBPF's capabilities from a security perspective. 

Despite the limitations and vulnerabilities noted in this application, the intership has provided an unique opportunity to engage with real world problems and leverage recent technologies to contribute to the advancement of knowledge in this field. This experience has also contributed to a greater appreciation for the importance of rigourous testing and evaluation in the development cycle.

In conclusion, while the application developed may not have achieved its inteded goals, the internship and the opportunity to conduct research on cutting-edge techonlogies proved to be an invaluable learning experience that has broadened the understanding of recent technologies and the potential applications in the field of security. The knowledge and skill will surely prove invaluable in future projects.


